### 2024-02-08 Main Paper/paper.qmd
Basic files and data of the paper

---
title: "Exploring Mate Competition As A Driver OF Anti Refugee Hate Crime"
author: 
  - Raghav Kanda
thanks: "Code and data are available at: [https://github.com/o8RaV/Exploring-Mate-Competition-as-a-Driver-of-Anti-Refugee-Hate-Crime/tree/main) [@knitr]"

date: Feburary 10, 2024
date-format: long
abstract: "The dataset obtained from the City of Toronto's OpenDataToronto Library sheds light on poll engagement dynamics within the city. With records starting from April 1, 2015, this study explores various facets of poll participation, including ballot distribution, cast and blank ballots, declarations added, and final voter counts."
format: pdf
number-sections: true
bibliography: references.bib
---

# Table of Contents

- Table of Contents

- Introduction

- Data
    - Source
    - Methodolgy
    - Features

- Results

- Discussion
    - Findings
    - Limitations
    - Future Research

- References

\newpage

# Introduction

\newpage

# Data

\newpage

# Results

\newpage

# Discussions

\newpage

# References
```{bibliography}
```