### 2024-02-09 table1.tex

% Table created by stargazer v.5.2.3 by Marek Hlavac, Social Policy Institute. E-mail: marek.hlavac at gmail.com
% Date and time: Fri, Feb 09, 2024 - 03:59:17 PM
\begin{table}[!htbp] \centering 
  \caption{} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{5}{c}{hate\_violence\_means} & formula.6 \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
\hline \\[-1.8ex] 
 Mate Competition & 0.437$^{**}$ & 0.263$^{**}$ & 0.236$^{**}$ & 0.206$^{**}$ & 0.185$^{**}$ & 0.155$^{**}$ \\ 
  & (0.016) & (0.020) & (0.021) & (0.019) & (0.019) & (0.019) \\ 
  & & & & & & \\ 
 Job Competition &  & 0.250$^{**}$ & 0.236$^{**}$ & 0.077$^{**}$ & 0.065$^{**}$ & 0.056$^{**}$ \\ 
  &  & (0.019) & (0.019) & (0.020) & (0.020) & (0.019) \\ 
  & & & & & & \\ 
 Life Satisfaction &  & $-$0.015$^{*}$ & $-$0.014$^{\dagger}$ & $-$0.003 & $-$0.002 & $-$0.0001 \\ 
  &  & (0.006) & (0.007) & (0.006) & (0.006) & (0.006) \\ 
  & & & & & & \\ 
\hline \\[-1.8ex] 
Observations & 3,019 & 3,019 & 3,008 & 3,008 & 3,008 & 3,008 \\ 
R$^{2}$ & 0.191 & 0.240 & 0.288 & 0.394 & 0.410 & 0.459 \\ 
Adjusted R$^{2}$ & 0.191 & 0.240 & 0.267 & 0.371 & 0.382 & 0.431 \\ 
Residual Std. Error & 0.799 (df = 3017) & 0.775 (df = 3015) & 0.760 (df = 2921) & 0.704 (df = 2897) & 0.698 (df = 2873) & 0.670 (df = 2857) \\ 
F Statistic & 714.588$^{**}$ (df = 1; 3017) & 317.891$^{**}$ (df = 3; 3015) & 13.756$^{**}$ (df = 86; 2921) & 17.141$^{**}$ (df = 110; 2897) & 14.878$^{**}$ (df = 134; 2873) & 16.170$^{**}$ (df = 150; 2857) \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{r}{$^{\dagger}$ p$<$0.1; $^{*}$ p$<$0.05; $^{**}$ p$<$0.01} \\ 
\end{tabular} 
\end{table}