### 2024-02-11 table1.tex

\documentclass{article}
\usepackage{booktabs} % for better-looking tables

\usepackage{geometry}

% Set your desired margins here
\geometry{
    left=1in,
    right=1in,
    top=1in,
    bottom=1in,
    % You can also set other margin parameters such as headheight, headsep, footskip, etc.
}

\begin{document}

\begin{table}[htbp]
\begin{tabular}{@{}|l|l|l|l|l|l|l|@{}}
\toprule
                                   & Model 1                                                     & Model 2                                                     & Model 3                                                     & Model 4                                                     & Model 5                                                     & Model 6                                                     \\ \midrule
Mate Competition                   & \begin{tabular}[c]{@{}l@{}}0.437 ***\\ (0.016)\end{tabular} & \begin{tabular}[c]{@{}l@{}}0.263 ***\\ (0.020)\end{tabular} & \begin{tabular}[c]{@{}l@{}}0.236 ***\\ (0.021)\end{tabular} & \begin{tabular}[c]{@{}l@{}}0.206 ***\\ (0.019)\end{tabular} & \begin{tabular}[c]{@{}l@{}}0.185 ***\\ (0.019)\end{tabular} & \begin{tabular}[c]{@{}l@{}}0.155 ***\\ (0.019)\end{tabular} \\ \midrule
Job Competition                    &                                                             & \begin{tabular}[c]{@{}l@{}}0.250 ***\\ (0.019)\end{tabular} & \begin{tabular}[c]{@{}l@{}}0.236 ***\\ (0.019)\end{tabular} & \begin{tabular}[c]{@{}l@{}}0.077 ***\\ (0.020)\end{tabular} & \begin{tabular}[c]{@{}l@{}}0.065 ***\\ (0.020)\end{tabular} & \begin{tabular}[c]{@{}l@{}}0.056 ***\\ (0.019)\end{tabular} \\ \midrule
Life Satisfaction                  &                                                             & \begin{tabular}[c]{@{}l@{}}-0.015*\\ (0.006)\end{tabular}   & \begin{tabular}[c]{@{}l@{}}-0.014\\ (0.007)\end{tabular}    & \begin{tabular}[c]{@{}l@{}}-0.003\\ (0.006)\end{tabular}    & \begin{tabular}[c]{@{}l@{}}-0.002\\ (0.006)\end{tabular}    & \begin{tabular}[c]{@{}l@{}}-0.0001\\ (0.006)\end{tabular}   \\ \midrule
Socio-Demographics                 &                                                             &                                                             & X                                                           & X                                                           & X                                                           & X                                                           \\ \midrule
National Attitudes Toward Refugees &                                                             &                                                             &                                                             & X                                                           & X                                                           & X                                                           \\ \midrule
Local Attitudes Toward Refugees    &                                                             &                                                             &                                                             &                                                             & X                                                           & X                                                           \\ \midrule
Additional Controls                &                                                             &                                                             &                                                             &                                                             &                                                             & X                                                           \\ \midrule
$N$                                & 3019                                                        & 3019                                                        & 3008                                                        & 3008                                                        & 3008                                                        & 3008                                                        \\ \midrule
$R^2$                              & 0.191                                                       & 0.240                                                       & 0.288                                                       & 0.394                                                       & 0.410                                                       & 0.459                                                       \\ \bottomrule
\end{tabular}
\end{table}
\end{document}